---
title: "Homework 01"
author: "Alec L. Robitaille"
date: "2023-02-01"
---


## Question 1
> Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 4
water and 11 land. Construct the posterior distribution.


```{r}
source('R/packages.R')
source('R/compute_posterior_globe.R')
```


```{r}
n_water <- 4
n_land <- 11
x <- c(rep('W', n_water), rep('L', n_land))
n_possibilities <- 5
posterior <- compute_posterior_globe(x, n_possibilities = n_possibilities)

kable(posterior, digits = 2)

# Using a finer grid with more potential possibilities
n_possibilities <- 100
posterior <- compute_posterior_globe(x, n_possibilities = n_possibilities)
```



## Question 2
> Using the posterior distribution from 1, compute the posterior predictive
distribution for the next 5 tosses of the same globe. I recommend you use the
sampling method.


```{r}
possibilities <- seq(0, 1, length.out = n_possibilities)
size <- 1e4
n_tosses <- 5


# == Sampling approach from posterior
# Sample with probability posterior distribution calculated in Question 1
# using the grid of possibilities (sequence of numbers between 0 and 1)
posterior_samples <- sample(
	possibilities,
	size = size,
	prob = posterior$post,
	replace = TRUE
)

# Using the posterior predicted proportion of water, simulate 5 tosses for each
posterior_predict <- rbinom(size, size = n_tosses, p = posterior_samples)

# Plot number of W tosses 
plot(table(posterior_predict), xlab = 'number of W tosses', ylab = '', 
		 main = 'Posterior predicted')

# === Sampling approach from representative Beta distribution
beta_samples <- rbeta(
	size,
	n_water + 1,
	n_land + 1
)

# Using the beta distribution predicted proportion of water, simulate 5 tosses for each
beta_predict <- rbinom(size, size = n_tosses, p = beta_samples)

# Plot number of W tosses 
plot(table(beta_predict), xlab = 'number of W tosses', ylab = '', 
		 main = 'Beta distribution predicted')
```


## Question 3
> Use the posterior predictive distribution from 2 to calculate the probability
of 3 or more water samples in the next 5 tosses.



## Question 4 (optional)
> This problem is an optional challenge for people who are taking the course for
a second or third time. Suppose you observe W = 5 water points, but you forgot
to write down how many times the globe was tossed, so you don’t know the number
of land points L. Assume that p = 0.7 and compute the posterior distribution of
the number of tosses N. Hint: Use the binomial distribution.
