---
title: "Homework 01"
author: "Alec L. Robitaille"
date: "2023-02-01"
---


## Question 1
> Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 4
water and 11 land. Construct the posterior distribution.


```{r}
source('R/packages.R')
source('R/compute_posterior_globe.R')
```


```{r}
n_water <- 4
n_land <- 11
x <- c(rep('W', n_water), rep('L', n_land))
n_possibilities <- 5
posterior <- compute_posterior_globe(x, n_possibilities = n_possibilities)

kable(posterior, digits = 2)

# Using a finer grid with more potential possibilities
n_possibilities <- 100
posterior <- compute_posterior_globe(x, n_possibilities = n_possibilities)
```



## Question 2
> Using the posterior distribution from 1, compute the posterior predictive
distribution for the next 5 tosses of the same globe. I recommend you use the
sampling method.


## Question 3
> Use the posterior predictive distribution from 2 to calculate the probability
of 3 or more water samples in the next 5 tosses.



## Question 4 (optional)
> This problem is an optional challenge for people who are taking the course for
a second or third time. Suppose you observe W = 5 water points, but you forgot
to write down how many times the globe was tossed, so you don’t know the number
of land points L. Assume that p = 0.7 and compute the posterior distribution of
the number of tosses N. Hint: Use the binomial distribution.
