---
title: "Homework 02"
author: "Alec L. Robitaille"
date: "2023-02-05"
---

```{r}
source('R/packages.R')
source('R/compute_posterior_globe.R')
```


## Question 1

> From the Howell1 dataset, consider only the people younger than 13 years
old. Estimate the causal association between age and weight. Assume that age
influences weight through two paths. First, age influences height, and height
influences weight. Second, age directly influences weight through age-related
changes in muscle growth and body proportions.

> Draw the DAG that represents these causal relationships. And then write a
generative simulation that takes age as an input and simulates height and
weight, obeying the relationships in the DAG.

```{r}
coords <- data.frame(
	name = c('H', 'A', 'W'),
	x =    c(1,    2,   3),
	y =    c(0,    1,   0)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dagify(
	W ~ H + A,
	H ~ A,
  coords = coords
) |> ggdag(seed = 2) + theme_dag()
```


```{r}
# Recall: https://speakerdeck.com/rmcelreath/statistical-rethinking-2023-lecture-04?slide=22


n_individuals <- 1e3
# Some numbers to guide the simulation
# 0 years: 50cm, 3.5kg
# 12 years: 150cm, 45kg

simulate_heights <- function(ages) {
	alpha_by_age <- seq(-2, 30, length.out = 12)
	height <- 50 + (ages * 10) + rnorm(length(ages), 0, 5)
	weight <- alpha_by_age[ages] + (0.1 * height) + rnorm(length(ages), 0, 0.3)
	
	return(data.table(
		age = ages,
		height = height,
		weight = weight
	))	
}


simulate_heights <- function(n) {
	bA <- 0.5
	bH <- 1
	
	A <- rnorm(n)
	H <- rnorm(n, bA * A)
	W <- rnorm(n, (bA * A)  + (bH * H))
	
	return(data.table(
		age = rescale(A, c(0, 12)),
		height = rescale(H, c(40, 200)),
		weight = rescale(W, c(2, 60))
	))	
}

DT <- simulate_heights(100)

m <- quap(data = DT, flist = alist(
	weight ~ dnorm(mu, sigma),
	mu <- alpha + beta_age * age,
	alpha ~ dnorm(3.5, 1),
	beta_age ~ dnorm(0, 5),
	sigma ~ dexp(1)
))

# Simulate n individuals with ages 0-12
DT <- simulate_heights(sample(seq.int(12), n_individuals, replace = TRUE))

```


## Question 2

> Use a linear regression to estimate the total causal effect of each year of
growth on weight.


```{r}


m <- quap()

```


## Question 3

> Now suppose the causal association between age and weight might be different
for boys and girls. Use a single linear regression, with a categorical variable
for sex, to estimate the total causal effect of age on weight separately for
boys and girls. How do girls and boys differ? Provide one or more posterior
contrasts as a summary.


## Question 4 - optional challenge

> The data in data(Oxboys) (rethinking package) are growth records for 26 boys
measured over 9 periods. I want you to model their growth. Specifically, model
the increments in growth from one period (Occasion in the data table) to the
next. Each increment is simply the difference between height in one occasion and
height in the pre- vious occasion. Since none of these boys shrunk during the
study, all of the growth increments are greater than zero. Estimate the
posterior distribution of these increments. Constrain the distribution so it is
always positive—it should not be possible for the model to think that boys can
shrink from year to year. Finally compute the posterior distribution of the
total growth over all 9 occasions.

