---
title: "Homework 02"
author: "Alec L. Robitaille"
date: "2023-02-05"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Question 1

> From the Howell1 dataset, consider only the people younger than 13 years
old. Estimate the causal association between age and weight. Assume that age
influences weight through two paths. First, age influences height, and height
influences weight. Second, age directly influences weight through age-related
changes in muscle growth and body proportions.

> Draw the DAG that represents these causal relationships. And then write a
generative simulation that takes age as an input and simulates height and
weight, obeying the relationships in the DAG.

```{r}
coords <- data.frame(
	name = c('H', 'A', 'W'),
	x =    c(1,    1,   2),
	y =    c(0,    1,   0)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dagify(
	W ~ H + A,
	H ~ A,
  coords = coords
) |> ggdag(seed = 2) + theme_dag()
```


```{r}
# Recall: https://speakerdeck.com/rmcelreath/statistical-rethinking-2023-lecture-04?slide=22

# Set number of individuals
n_individuals <- 1e3

# Some numbers to guide the simulation
# 0 years: 50cm, 3.5kg
# 12 years: 150cm, 45kg
simulate_heights <- function(n) {
	beta_age_height <- 7
	beta_height_weight <- 3
	beta_age_weight <- 2.5
	
	age <- runif(n, 0, 12)
	height <- rnorm(n, beta_age_height * age, 2)
	weight <- rnorm(n, 
									(beta_height_weight * height) + (beta_age_weight * age), 
									2)
	
	return(data.table(
		age = age,
		height = height,
		weight = weight
	))	
}


DT <- simulate_heights(20)

ggplot(DT) + 
	geom_point(aes(age, weight))
```

## Question 2

> Use a linear regression to estimate the total causal effect of each year of
growth on weight.


```{r}
source('R/data_Howell1.R')

# height: Height in cm
# weight: Weight in kg
# age: Age in years
DT <- data_Howell()[age < 12]

# Simulate to determine appropriate priors
n <- 100
ggplot() + 
	geom_abline(aes(intercept = rnorm(n, 4, 0.5), 
									slope = rnorm(n, 4, 1)),
							alpha = 0.1, size = 2) + 
	labs(x = 'age (cm)', y = 'weight (kg)') + 
	xlim(0, 12) + 
	ylim(0, 50)

# Model
m <- quap(
	alist(
		weight ~ dnorm(mu, sigma),
		mu <- alpha + beta_age * age,
		alpha ~ dnorm(4, 0.5),
		beta_age ~ dnorm(4, 1),
		sigma ~ dexp(1)
	),
	data = DT
)

precis(m)
```


## Question 3

> Now suppose the causal association between age and weight might be different
for boys and girls. Use a single linear regression, with a categorical variable
for sex, to estimate the total causal effect of age on weight separately for
boys and girls. How do girls and boys differ? Provide one or more posterior
contrasts as a summary.


```{r}
coords <- data.frame(
	name = c('H', 'A', 'W', 'S'),
	x =    c(1,    1,   2,    2),
	y =    c(0,    1,   0,    1)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dagify(
	W ~ H + A + S,
	H ~ A + S,
  coords = coords
) |> ggdag(seed = 2) + theme_dag()
```

```{r}
#| warning: false


# Group index for sex
DT[, sex := .GRP, by = male]

# Set up key of sex factor levels and labels
key <- unique(DT[, .(sex, sex_char = fifelse(male == 1, 'male', 'female'))])

# Model
m <- quap(
	alist(
		weight ~ dnorm(mu, sigma),
		mu <- alpha[sex] + beta_age[sex] * age,
		alpha[sex] ~ dnorm(4, 0.5),
		beta_age[sex] ~ dnorm(4, 1),
		sigma ~ dexp(1)
	),
	data = DT
)

# Set up the options for predictions
DT_pred <- data.table(
	age = seq(0, 12, 0.5)
)[, variable := paste0('V', .I)]


# Simulate where sex = 1, and where sex = 2
sim_sex_1 <- sim(m, DT_pred[, .(age, sex = 1)])
sim_sex_2 <- sim(m, DT_pred[, .(age, sex = 2)])

# Posterior of differences
sim_diff <- sim_sex_1 - sim_sex_2

# Melt and tidy
melt_diff <- melt(as.data.table(sim_diff))
DT_diff <- melt_diff[DT_pred, on = 'variable']


# Plot
ggplot(DT_diff, aes(age, value)) +
	stat_lineribbon(.width = c(0.5, 0.8, 0.95)) +
	scale_fill_scico_d(direction = -1,
										 palette = 'nuuk',
										 end = 0.9) +
	labs(
		y = 'predicted difference in weight',
		title = paste('Posterior predicted difference in weight between',
									key[sex == 1, sex_char], 'and',
									key[sex == 2, sex_char])
	)


# Note: differences with link
# Predict where sex = 1, and where sex = 2
link_sex_1 <- link(m, DT_pred[, .(age, sex = 1)])
link_sex_2 <- link(m, DT_pred[, .(age, sex = 2)])

# Posterior of differences
link_diff <- link_sex_1 - link_sex_2

# Melt and tidy
melt_diff <- melt(as.data.table(link_diff))
DT_diff_link <- melt_diff[DT_pred, on = 'variable']


# Plot
ggplot(DT_diff_link, aes(age, value)) +
	stat_lineribbon(.width = c(0.5, 0.8, 0.95)) +
	scale_fill_scico_d(direction = -1,
										 palette = 'nuuk',
										 end = 0.9) +
	labs(
		y = 'predicted difference in weight',
		title = paste('Predicted difference in weight between',
									key[sex == 1, sex_char], 'and',
									key[sex == 2, sex_char])
	)

```

Details on the difference between `sim` and `link` are in the
textbook around pages 105-110, the Overthinking box "How does link work?", 
code 4.58-4.62, Rethinking box "two kinds of uncertainty". 

- link takes the model formula to return what the average line would
predict about your new data
- sim takes the model formula and **uncertainty** to return
what the posterior distribution of lines would predict about your 
new data


## Question 4 - optional challenge

> The data in data(Oxboys) (rethinking package) are growth records for 26 boys
measured over 9 periods. I want you to model their growth. Specifically, model
the increments in growth from one period (Occasion in the data table) to the
next. Each increment is simply the difference between height in one occasion and
height in the previous occasion. Since none of these boys shrunk during the
study, all of the growth increments are greater than zero. Estimate the
posterior distribution of these increments. Constrain the distribution so it is
always positive — it should not be possible for the model to think that boys can
shrink from year to year. Finally compute the posterior distribution of the
total growth over all 9 occasions.


```{r}
# Get data and prepare it
data(Oxboys)

DT_ox <- data.table(Oxboys)

setorder(DT_ox, Occasion)
setnames(DT_ox, colnames(DT_ox), make_clean_names(colnames(DT_ox)))
DT_ox[, diff_height := height - shift(height), by = subject]
DT_ox[, index_occasion := occasion - 1]


# Simulate to evaluate priors
n <- 100
sim_prior <- CJ(occasion = unique(DT_ox$occasion),
								row_id = seq.int(n))
sim_prior[, diff_height := rexp(.N, 1)]

ggplot(sim_prior, aes(factor(occasion), diff_height)) +
	stat_slab() +
	labs(x = 'ocassion', y = 'difference in height (cm)')


m <- quap(
	alist(
		diff_height ~ dnorm(mu, sigma),
		mu <- alpha[index_occasion],
		alpha[index_occasion] ~ dexp(1),
		sigma ~ dexp(1)
	),
	data = DT_ox[!is.na(diff_height)]
)

plot(precis(m, 2))


samples <- extract.samples(m)$alpha

melt_samples <- melt(as.data.table(samples))
melt_samples[, index_occasion := as.integer(gsub('V', '', variable))]

sims <- sim(m, data.table(index_occasion = seq.int(8)))
melt_sim <- melt(as.data.table(sims))
melt_sim[, index_occasion := as.integer(gsub('V', '', variable))]

ggplot(melt_samples, aes(factor(index_occasion + 1, levels = seq.int(8) + 1), value)) + 
	stat_slab() + 
	stat_dots(aes(y = height_diff), data = DT_ox, color = 'black', alpha = 0.2) + 
	stat_slab(data = melt_sim, color = 'blue', alpha = 0.2, fill = 'lightblue') + 
	labs(x = 'occasion', y = 'height difference')

# Posterior distribution of total growth 
dens(apply(sims, 1, sum))
```
