---
title: "Homework 03"
author: "Alec L. Robitaille"
date: "2023-03-01"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 
```{r}
# Data
source('R/data_foxes.R')

# group : ID of group
# avgfood : Average available food in group's territory
# groupsize : Size of each group
# area : Area of group territory
# weight : Body weight of individual fox
DT <- data_foxes(scale = TRUE)
```


## Question 1
> The first two problems are based on the same data. The data in data(foxes) are
116 foxes from 30 different urban groups in England. These fox groups are like
street gangs. Group size (groupsize) varies from 2 to 8 individuals. Each group
maintains its own (almost exclusive) urban territory. Some ter- ritories are
larger than others. The area variable encodes this information. Some territories
also have more avgfood than others. And food influences the weight of each fox.
Assume this following DAG. Use the backdoor criterion and estimate the total
causal influence of A on F. What effect would increasing the area of a territory
have on the amount of food inside it?


### DAG
```{r}
coords <- data.frame(
	name = c('A', 'F', 'G', 'W'),
	x =    c(2,    1,   3,    2),
	y =    c(3,    2,   2,    1)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dag <- dagify(
	W ~ F + G,
	F ~ A,
	G ~ F,
  coords = coords
)
dag |> ggdag(seed = 2) + theme_dag()
```

- A: area
- F: avgfood
- G: groupsize
- W: weight



### Estimand

Total causal influence of area on average food. Consider how an intervention
on area would influence average food. 


```{r}
adjustmentSets(dag, exposure = 'A', outcome = 'F', effect = 'total')
```


### Simulation

We don't know much about this data, so to just get an idea of the units
of the data, check the range of each column.

```{r}
lapply(DT, range)
```


But instead of peaking, let's use the standardized versions and consider the 
potential effects on the scale of standard deviations. 

```{r}
#| warning: false
tar_load(m_h01_q01_prior)

m_h01_q01_prior$formula
m_h01_q01_prior$prior

# Show draws from prior distributions
conditional_effects(m_h01_q01_prior, 'scale_area', prob = 0.89)
```

### Model

```{r}
#| warning: false
tar_load(m_h01_q01)

m_h01_q01$formula
m_h01_q01$prior

mcmc_areas(m_h01_q01, pars = c('b_Intercept', 'b_scale_area', 'sigma'))

# Total causal influence of area on average food 
conditional_effects(m_h01_q01, 'scale_area', prob = 0.89)
```


## Question 2

> Infer the total causal effect of adding food F to a territory on the weight W of
foxes. Can you calculate the causal effect by simulating an intervention on
food?


### Estimand

Total causal influence of average food on weight. Consider how an intervention
on average food would influence weight. 


```{r}
adjustmentSets(dag, exposure = 'F', outcome = 'W', effect = 'total')
```



### Simulation

```{r}
#| warning: false
tar_load(m_h01_q02_prior)

m_h01_q02_prior$formula
m_h01_q02_prior$prior

# Show draws from prior distributions
conditional_effects(m_h01_q02_prior, 'scale_avgfood', prob = 0.89)
```

### Model

```{r}
#| warning: false
tar_load(m_h01_q02)

m_h01_q02$formula
m_h01_q02$prior

mcmc_areas(m_h01_q01, pars = c('b_Intercept', 'b_scale_avgfood', 'sigma'))

# Total causal influence of average food on weight
conditional_effects(m_h01_q02, 'scale_avgfood', prob = 0.89)
```



## Question 3

> Infer the direct causal effect of adding food F to a territory on the weight W
of foxes. In light of your estimates from this problem and the previous one,
what do you think is going on with these foxes?


### Estimand

Direct causal influence of average food on weight. Consider how an intervention
on average food would influence weight. 


```{r}
adjustmentSets(dag, exposure = 'F', outcome = 'W', effect = 'direct')
```



### Simulation

```{r}
#| warning: false
tar_load(m_h01_q03_prior)

m_h01_q03_prior$formula
m_h01_q03_prior$prior

# Show draws from prior distributions
conditional_effects(m_h01_q03_prior, 'scale_avgfood', prob = 0.89)
```

### Model

```{r}
#| warning: false
tar_load(m_h01_q03)

m_h01_q03$formula
m_h01_q03$prior

mcmc_areas(m_h01_q01, pars = c('b_Intercept', 'b_scale_avgfood', 'sigma'))

# Direct causal influence of average food on weight
conditional_effects(m_h01_q03, 'scale_avgfood', prob = 0.89)
```

Closing the backdoor path and estimating the direct effect of average food
on weight now shows a positive relationship between average food and fox weight. 


```{r}
conditional_effects(m_h01_q03, 'scale_groupsize', prob = 0.89)

#| warning: false
tar_load(m_h01_q03_groupsize_food)

m_h01_q03_groupsize_food$formula
m_h01_q03_groupsize_food$prior

# Direct causal influence of average food on group size
conditional_effects(m_h01_q03_groupsize_food, 'scale_avgfood', prob = 0.89)
```

As average food in a territory increases, the group size increases. 



## Question 4 (bonus)

> Suppose there is an unobserved confound that influences F and G.
Assuming the DAG above is correct, again estimate both the total and
direct causal effects of F on W. What impact does the unobserved confound have?


```{r}
coords <- data.frame(
	name = c('A', 'F', 'G', 'W', 'U'),
	x =    c(1,    1,   2,    1.5,   2),
	y =    c(3,    2,   2,    1,   3)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dag <- dagify(
	W ~ F + G,
	F ~ A + U,
	G ~ F + U,
  coords = coords
)
dag |> ggdag(seed = 2) + theme_dag()
```

### Estimand

Total and direct causal effects of F on W. 


```{r}
adjustmentSets(dag, exposure = 'F', outcome = 'W', effect = 'total')
adjustmentSets(dag, exposure = 'F', outcome = 'W', effect = 'direct')
```

In estimating the total causal effect of average food on weight, the 
unobserved confound U opens a backdoor on average food through 
group size. If group size is included, the direct effect will be estimated
and not the total effect. 
