---
title: "Homework 03"
author: "Alec L. Robitaille"
date: "2023-03-01"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 
```{r}
# Data
source('R/data_foxes.R')

# group : ID of group
# avgfood : Average available food in group's territory
# groupsize : Size of each group
# area : Area of group territory
# weight : Body weight of individual fox
DT <- data_foxes(scale = TRUE)
```


## Question 1
> The first two problems are based on the same data. The data in data(foxes) are
116 foxes from 30 different urban groups in England. These fox groups are like
street gangs. Group size (groupsize) varies from 2 to 8 individuals. Each group
maintains its own (almost exclusive) urban territory. Some ter- ritories are
larger than others. The area variable encodes this information. Some territories
also have more avgfood than others. And food influences the weight of each fox.
Assume this following DAG. Use the backdoor criterion and estimate the total
causal influence of A on F. What effect would increasing the area of a territory
have on the amount of food inside it?


### DAG
```{r}
coords <- data.frame(
	name = c('A', 'F', 'G', 'W'),
	x =    c(2,    1,   3,    2),
	y =    c(3,    2,   2,    1)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dag <- dagify(
	W ~ F + G,
	F ~ A,
	G ~ F,
  coords = coords
)
dag |> ggdag(seed = 2) + theme_dag()
```

- A: area
- F: avgfood
- G: groupsize
- W: weight



### Estimand

Total causal influence of area on average food. Consider how an intervention
on area would influence average food. 


```{r}
adjustmentSets(dag, exposure = 'A', outcome = 'F', effect = 'total')
```


### Simulation

We don't know much about this data, so to just get an idea of the units
of the data, check the range of each column.

```{r}
lapply(DT, range)
```


But instead of peaking, let's use the standardized versions and consider the 
potential effects on the scale of standard deviations. 

```{r}
#| warning: false
tar_load(m_h01_q01_prior)

m_h01_q01_prior$formula
m_h01_q01_prior$prior

# Show draws from prior distributions
conditional_effects(m_h01_q01_prior, 'scale_area', prob = 0.89)
```

### Model

```{r}
#| warning: false
tar_load(m_h01_q01)

m_h01_q01$formula
m_h01_q01$prior

# Total causal influence of area on average food 
conditional_effects(m_h01_q01, 'scale_area', prob = 0.89)
```


## Question 2

> Infer the total causal effect of adding food F to a territory on the weight W of
foxes. Can you calculate the causal effect by simulating an intervention on
food?


### Estimand

Total causal influence of average food on weight. Consider how an intervention
on average food would influence weight. 


```{r}
adjustmentSets(dag, exposure = 'F', outcome = 'W', effect = 'total')
```



### Simulation

```{r}
#| warning: false
tar_load(m_h01_q02_prior)

m_h01_q02_prior$formula
m_h01_q02_prior$prior

# Show draws from prior distributions
conditional_effects(m_h01_q02_prior, 'scale_avgfood', prob = 0.89)
```

### Model

```{r}
#| warning: false
tar_load(m_h01_q02)

m_h01_q02$formula
m_h01_q02$prior

# Total causal influence of average food on weight
conditional_effects(m_h01_q02, 'scale_avgfood', prob = 0.89)
```
