---
title: "Homework 05"
author: "Alec L. Robitaille"
date: "2023-04-28"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 
```{r}
# Data
source('R/data_grants.R')

# discipline
# gender
# applications
# awards

DT_grants <- data_grants()
```


## Question 1 

> The data in data(NWOGrants) are outcomes for scientific funding applications
for the Netherlands Organization for Scientific Research (NWO) from 2010–2012
(see van der Lee and Ellemers doi:10.1073/pnas.1510159112). These data have a
structure similar to the UCBAdmit data discussed in Chapter 11 and in lecture.
There are applications and each has an associated gender (of the lead researcher).
But instead of departments, there are disciplines. Draw a DAG for this sample. Then
use the backdoor criterion and a binomial GLM to estimate the TOTAL causal effect
of gender on grant awards


```{r h05_q01_dag}
coords <- data.frame(
    name = c('G', 'D', 'A'),
    x =    c(1,    2,   3),
    y =    c(0,    1,   0)
)

dagify(
	A ~ G + D,
	D ~ G + A,
	coords = coords
) |> ggdag(seed = 2) + theme_dag()
```

Total effect of gender on award has two paths: through discipline and directly to award. 

Prior predictive simulation

```{r}
get_prior(
    awards | trials(applications) ~ gender,
    family = 'binomial',
    data = DT_grants
)
```

```{r h05_q01_prior}
tar_load(m_h05_q01_prior)
m_h05_q01_prior$priors

mcmc_areas(m_h05_q01_prior, regex_pars = 'b')
```


Posterior

```{r h05_q01_posterior}
tar_load(m_h05_q01)
m_h05_q01

mcmc_areas(m_h05_q01, regex_pars = 'b')
```

Estimated marginal means

```{r h05_q01_emmean}
emmean_total_gender <- emmeans(m_h05_q01, ~ gender, regrid = 'response')
emmean_total_gender
plot(emmean_total_gender, level = .87) + xlim(0, 1) + xlab('award')
```

Contrast

```{r} 
contrast(emmean_total_gender, method = 'pairwise')
```


## Question 2

> Now estimate the DIRECT causal effect of gender on grant awards. Use the same
DAG as above to justify one or more binomial models. Compute the average direct
causal effect of gender, weighting each discipline in proportion to the number of
applications in the sample. Refer to the marginal effect example in Lecture 9 for
help.

Direct effect of gender on award requires conditioning on discipline. 


Posterior

```{r h05_q02_posterior}
tar_load(m_h05_q02)
m_h05_q02

mcmc_areas(m_h05_q02, regex_pars = 'b')
```

Estimated marginal means

```{r h05_q02_emmean}
emmean_direct_gender <- emmeans(m_h05_q02, ~ gender, regrid = 'response')
emmean_direct_gender
plot(emmean_direct_gender, level = .87) + xlim(0, 1) + xlab('award')
```

Contrast

```{r}
contrast(emmean_direct_gender, method = 'pairwise')
```

