---
title: "Homework 05"
author: "Alec L. Robitaille"
date: "2023-04-28"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 
```{r}
# Data
source('R/data_grants.R')

# discipline
# gender
# applications
# awards

DT_grants <- data_grants()
```


## Question 1 

> The data in data(NWOGrants) are outcomes for scientific funding applications
for the Netherlands Organization for Scientific Research (NWO) from 2010–2012
(see van der Lee and Ellemers doi:10.1073/pnas.1510159112). These data have a
structure similar to the UCBAdmit data discussed in Chapter 11 and in lecture.
There are applications and each has an associated gender (of the lead researcher).
But instead of departments, there are disciplines. Draw a DAG for this sample. Then
use the backdoor criterion and a binomial GLM to estimate the TOTAL causal effect
of gender on grant awards


```{r}
coords <- data.frame(
    name = c('G', 'D', 'A'),
    x =    c(1,    2,   3),
    y =    c(0,    1,   0)
)

dagify(
	A ~ G + D,
	D ~ G + A,
	coords = coords
) |> ggdag(seed = 2) + theme_dag()
```

Total effect of gender on award has two paths: through discipline and directly to award. 

Prior predictive simulation 

```{r}
tar_load(m_h05_q01_prior)

conditional_effects(m_h05_q01_prior)
```