---
title: "Homework 06"
author: "Alec L. Robitaille"
date: "2023-04-26"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 
```{r}
# Data
source('R/data_reedfrogs.R')

# from ?reedfrogs
# density: initial tadpole density (number of tadpoles in a 1.2 x 0.8 x 0.4 m tank) [experiment 1]
# pred: factor -  predators present or absent [experiment 1]
# size: factor - big or small tadpoles [experiment 1]
# surv: number surviving
# propsurv: proportion surviving (=surv/density) [experiment 1]

DT_frogs <- data_reedfrogs()
```


## Question 1
> Conduct a prior predictive simulation for the Reedfrog model. By this I mean
to simulate the prior distribution of tank survival probabilities αj. Start by
using this prior: 

> $\alpha_{j} \sim Normal(\bar{\alpha}, \sigma)$

> $\bar{\alpha} \sim Normal(0, 1)$

> $\sigma \sim Exponential(1)$ 

> Be sure to transform the $\alpha_{j}$ values to the probability scale for
plotting and summary. How does increasing the width of the prior on $\sigma$
change the prior distribution of $\alpha_{j}$? You might try Exponential(10) and
Exponential(0.1) for example.



```{r}
get_prior(
	surv | trials(density) ~ (1 | tank),
	data = DT_frogs,
	family = 'binomial'
)
```

- $\bar{\alpha}$ corresponds to class Intercept
- $\sigma$ corresponds to class sd


From: Bürkner, P.-C. (2017). brms: An R Package for Bayesian Multilevel Models
Using Stan. Journal of Statistical Software, 80(1), 1–28.
https://doi.org/10.18637/jss.v080.i01

> `fit1 <- brm(formula = time | cens(censored) ~ age * sex + disease + (1 +
age|patient), data = kidney, family = lognormal(), prior =
c(set_prior("normal(0,5)", class = "b"), set_prior("cauchy(0,2)", class = "sd"),
set_prior("lkj(2)", class = "cor")))`
  
> Each group-level effect of each grouping factor has a standard deviation
parameter, which is restricted to be non-negative ... In brms, standard
deviation parameters are named as sd_<group>_<coef> so that sd_patient_Intercept
and sd_patient_age are the parameter names in the example [(1 + age|patient)].



::: {.panel-tabset}

## Exponential(1)
```{r}
tar_load(m_h06_q01_prior_exp_1)
m_h06_q01_prior_exp_1$prior
```


```{r prior_exp_1_areas}
#| fig-width: 7
#| fig-height: 12
mcmc_areas(
	m_h06_q01_prior_exp_1,
	regex_pars = 'r_tank', 
	transformations = inv_logit
)
```

## Exponential(0.1)

```{r}
tar_load(m_h06_q01_prior_exp_0pt1)
m_h06_q01_prior_exp_0pt1$prior
```


```{r prior_exp_0pt1_areas}
#| fig-width: 7
#| fig-height: 12
mcmc_areas(
	m_h06_q01_prior_exp_0pt1,
	regex_pars = 'r_tank',
	transformations = inv_logit
)
```

## Exponential(10)

```{r}
tar_load(m_h06_q01_prior_exp_10)
m_h06_q01_prior_exp_10$prior
```


```{r prior_exp_10_areas}
#| fig-width: 7
#| fig-height: 12
mcmc_areas(
	m_h06_q01_prior_exp_10, 
	regex_pars = 'r_tank', 
	transformations = inv_logit
)
```
:::




## Question 2

> Revisit the Reedfrog survival data, data(reedfrogs). Start with the varying
effects model from the book and lecture. Then modify it to estimate the causal
effects of the treatment variables pred and size, including how size might
modify the effect of predation. An easy approach is to estimate an effect for
each combination of pred and size. Justify your model with a DAG of this
experiment.


```{r}
coords <- data.frame(
    name = c('D', 'G', 'P', 'T', 'S'),
    x =    c(1,    2,   3,   1,    2),
    y =    c(0,    0,   0,   1,  1)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4
dagify(
    S ~ D + G + P + T,
  coords = coords
) |> ggdag(seed = 2, layout = 'auto') + theme_dag()
```

Survival in a tank depends on density, tank size, and predation. 

$S_{i} \sim Binomial(D_{i}, p_{i} )$

$logit(p_{i}) = \alpha_{T[i]} + \beta_{P, S}P_{i}$

$\alpha_{j} \sim Normal(\bar{\alpha}, \sigma)$

$\bar{\alpha} \sim Normal(0, 1)$

$\beta_{P, S} \sim Normal(0, 0.5)$

$\sigma \sim Exponential(1)$ 


```{r}
get_prior(
	surv | trials(density) ~ pred * size + (1 | tank),
	data = DT_frogs,
	family = 'binomial'
)
```

- $\bar{\alpha}$ corresponds to class Intercept
- $\sigma$ corresponds to class sd


```{r}
tar_load(m_h06_q02)
m_h06_q02
m_h06_q02$prior
```

Reminder: 

- Prediction includes uncertainty of coefficients, variance parameters of the groups and each individual observation
- Epred (expected values) includes the uncertainty of the coefficients and variance parameters of the groups
- Emmeans returns the average marginal effects



```{r, h06_q02_emmean}
# Average marginal effects
emmean_pred <- emmeans(
	m_h06_q02, 
	~ pred * size, 
	regrid = 'response',
	at = list(density = c(25))
) |> gather_emmeans_draws()

ggplot(emmean_pred) + 
	stat_halfeye(aes(.value, size, fill = pred), alpha = 0.7) + 
	labs(x = 'Survival', y = 'Size')  +
	scale_fill_scico_d(begin = 0.3)
```

Tanks with predation had lower survival, and smaller tanks had higher survival than bigger tanks. There is more uncertainty in tanks with predation.


## Question 3

> Now estimate the causal effect of density on survival. Consider whether pred
modifies the effect of density. There are several good ways to include density
in your Binomial GLM. You could treat it as a continuous regression variable
(possibly standardized). Or you could convert it to an ordered category (with
three levels). Compare the σ (tank standard deviation) posterior distribution
to σ from your model in Problem 2. How are they different? Why?

```{r}
tar_load(m_h06_q03)
m_h06_q03
m_h06_q03$prior
```

```{r, h06_q03_emmean}
# Average marginal effects
emmean_pred_dens <- emmeans(
	m_h06_q03, 
	~ pred * density, 
	regrid = 'response',
	at = list(density = c(10, 25, 35))
) |> gather_emmeans_draws()

ggplot(emmean_pred_dens) +
	stat_halfeye(aes(.value, factor(density), fill = pred), alpha = 0.7) +
	labs(x = 'Survival', y = 'Initial tank density') +
	scale_fill_scico_d(begin = 0.3)
```



Compare the \sigma from Question 2

```{r}
(mcmc_areas(m_h06_q02, 'sd_tank__Intercept') + xlim(0, 2) + labs(title = 'Question 2')) / 
	(mcmc_areas(m_h06_q03, 'sd_tank__Intercept') + xlim(0, 2) + labs(title = 'Question 3'))
```

The distribution for tank standard deviation for Question 3 has less density at
the upper ranges (>1.25) compared to the tank standard deviation for Question 2.
The tank standard deviation is reduced because some of the variability between
tanks was due to density and now that it is included in the model, this
variability is incorporated into that parameter.
