---
title: "Homework 06"
author: "Alec L. Robitaille"
date: "2023-04-26"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 
```{r}
# Data
source('R/data_reedfrogs.R')

# from ?reedfrogs
# density: initial tadpole density (number of tadpoles in a 1.2 x 0.8 x 0.4 m tank) [experiment 1]
# pred: factor -  predators present or absent [experiment 1]
# size: factor - big or small tadpoles [experiment 1]
# surv: number surviving
# propsurv: proportion surviving (=surv/density) [experiment 1]

DT <- data_reedfrogs()
```


## Question 1
> Conduct a prior predictive simulation for the Reedfrog model. By this I mean
to simulate the prior distribution of tank survival probabilities αj. Start by
using this prior: 

> $\alpha_{j} \sim Normal(\bar{\alpha}, \sigma)$

> $\alpha \sim Normal(0, 1)$

> $\sigma \sim Exponential(1)$ 

> Be sure to transform the $\alpha_{j}$ values to the probability scale for
plotting and summary. How does increasing the width of the prior on $\sigma$ change the
prior distribution of $\alpha_{j}$? You might try Exponential(10) and
Exponential(0.1) for example.
