---
title: "Homework 09"
author: "Alec L. Robitaille"
date: "2023-05-04"
---

```{r, include = FALSE}
source('R/packages.R')
```

```{r, eval = FALSE}
source('R/packages.R')
```


## Data 

The data in data(Achehunting) are 14,364 individual hunting trips by 147 men
among the indigenous Aché of Paraguay. Each trip has recorded the hunter’s identity
(anonymized id), hunter’s age at the time of the trip, the duration in hours of the
trip, and the kilograms of meat returned. Many of the trips have missing values for
duration.

```{r}
# Data
source('R/data_achehunting.R')

# month : Month of record
# day : Day of record
# year : Year of record
# id : Identifier of individual man
# age : Man's age at time of record
# kg.meat : Kilograms of meat returned from hunt
# hours : Duration in hours of hunting trip
# datatype : 1 if duration of trip known, 3 otherwise

DT_hunting <- data_achehunting()
```



## Question 1

> 1. In this first problem, estimate the influence of age on the probability of trip success. Define “success” as a trip that returns any non-zero amount of meat. Ignore
individual hunter identities for now. You can use any functional relationship for age
and success that you think is sensible, but be sure to justify it and check it against
posterior predictions.


```{r h09_dag_q01}
coords <- data.frame(
    name = c('S', 'A'),
    x =    c(1,    2),
    y =    c(0,    1)
)

dagify(
	S ~ A,
	coords = coords
) |> ggdag(seed = 2) + theme_dag()
```

### Functional relationship

First, fit with a log base 10 line. 

```{r m_h09_q01_function}
range_age <- data.table(
    age = seq(min(DT_hunting$age), max(DT_hunting$age))
)

ggplot(DT_hunting[, .(prob_success = sum(success) / .N), age], 
    aes(x = age, prob_success)) + 
    geom_point() + 
    stat_function(fun = function(x) -1 + log10(x)) + 
    labs(x = 'Age', y = 'Probability of success')
```

Short of more background on these data, I am going to take it as practice fitting
a GAM with `brms` (through `mgcv`).


### Priors
```{r}
get_prior(success ~ s(age), family = 'binomial', data = DT_hunting)
```


Recall: predicted draws returns posterior predictions on the scale of the response 
(0, 1), epred draws returns posterior predictions with the inverse transformation. 

```{r h09_q01_prior}
tar_load(m_h09_q01_prior)
m_h09_q01_prior$prior

pred_q01_prior <- add_predicted_draws(range_age, m_h09_q01_prior)
table(pred_q01_prior$.prediction)

epred_q01_prior <- add_epred_draws(range_age, m_h09_q01_prior)
range(epred_q01_prior$.epred)

ggplot(epred_q01_prior) + 
    stat_lineribbon(aes(age, .epred)) + 
    ylim(0, 1) +
    scale_fill_grey(start = 0.8, end = 0.4) + 
    labs(x = 'Age', y = 'Prior probability of success')
```

### Posterior

```{r h09_q01}
tar_load(m_h09_q01)
m_h09_q01

epred_q01 <- add_epred_draws(range_age, m_h09_q01)
range(epred_q01$.epred)

ggplot(epred_q01) + 
    stat_lineribbon(aes(age, .epred)) + 
    ylim(0, 1) +
    scale_fill_grey(start = 0.8, end = 0.4) + 
    labs(x = 'Age', y = 'Posterior probability of success') + 
    geom_point(aes(age, prob_success), 
               data = DT_hunting[, .(prob_success = sum(success) / .N), age])
```



## Question 2

> 2. Now incorporate individual hunter varying effects into your analysis from the
previous problem. Allow the influence of age on success to vary by each individual.
How much variation in success is explained by individuals and how much by age?

### Priors

```{r}
get_prior(success ~ age + (1 | id), family = 'binomial', data = DT_hunting)
```

### Posterior

```{r}
tar_load(m_h09_q02)
m_h09_q02

mcmc_areas(m_h09_q02, regex_pars = 'sd|b')
```

## Question 3

> 3. Now include trip duration in the model from the previous problem. The focus
is still the causal relationship between age and success, but duration is a potential
competing cause or a mediator, so it can be useful to include it for either increased
efficiency (competing cause) or to study how variation in trip durations mediate differences in hunting returns. Since there are many missing values for trip duration,
compare (1) a complete case analysis, which drops all the trips with missing durations, to (2) an analysis which uses all 14,364 trips and imputes missing duration
values where necessary. Use any functional relationship between duration and success that you think is sensible, but be sure to justify it and check it against posterior
predictions.

```{r}
tar_load(m_h09_q03_complete_cases)
m_h09_q03_complete_cases

mcmc_areas(m_h09_q03_complete_cases, regex_pars = 'sd|b')
conditional_effects(m_h09_q03_complete_cases)
```





```{r}

imp <- tar_read(DT_hunting_mice)



tar_load(m_h09_q03_mice)
m_h09_q03_mice

mcmc_areas(m_h09_q03_mice, regex_pars = 'sd|b')

conditional_effects(m_h09_q03_mice)
```