---
title: "Lecture 12 Notes"
author: "Alec L. Robitaille"
date: "2023-02-10"
---

```{r, include = FALSE}
source('R/packages.R')
```


## Multilevel models

Repeat observations within groups can be modelled using categorical variables eg.
a vector of alphas for each individual. Categorical variables have 
*anterograde amnesia*, when estimating eg. the beta for a new individual, 
this estimate does not contribute anything to the next individual. 

Multilevel models instead, have two kinds of models

1. Model observations (eg. responses) of groups/individuals
1. Model population of groups/individuals (clusters)

The population model creates a kind of memory. Advantages of multilevel model
include: 

- more efficient estimation 
- resists overfitting

The order that we estimate clusters does not matter, because we update the 
population level model with every cluster estimate and that 
population level model contributes to all clusters' estimates. The 
information gathered from each cluster is pooled. 


### Regularization

Types of pooling

- Complete pooling: all clusters are the same. This results in underfitting. 
- No pooling: all clusters are unrelated. This results in overfitting. 
- Partial pooling: adaptive compromise. 


Regularization with multilevel models is partial pooling, an adaptive 
compromise. 


### Example: cafes

Grey distribution is the priors
Red distribution is the posterior
Blue distribution is the population

### Example: reedfrogs

- 48 tanks (T) of reedfrogs
- treatments: density (D), size (G), predation (P)
- outcome: survival (S)


```{r}
coords <- data.frame(
	name = c('D', 'G', 'P', 'T', 'S'),
	x =    c(1,    2,   3,   1,    2),
	y =    c(0,    0,   0,   1,  1)
)
```

```{r}
#| fig.height: 4
#| fig.width: 4 
dagify(
	S ~ D + G + P + T,
  coords = coords
) |> ggdag(seed = 2, layout = 'auto') + theme_dag()
```

### Example: repeat observations in the trolley data

Variation in the responses across stories and across individual participants. 

